---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(data.table)
unzip("./activity.zip")
activity <- fread("./activity.csv")
```



## What is mean total number of steps taken per day?
```{r results='hide'}
stepsPerDay <- activity[!is.na(steps)][,.(steps=sum(steps)),by=date]
totalSteps <- sum(stepsPerDay$steps)
meanSteps <- mean(stepsPerDay$steps)
medianSteps <- median(stepsPerDay$steps)
hist(stepsPerDay$steps)
```

Total number of steps: `r totalSteps`

Mean number of steps per day: `r meanSteps`

Median number of steps per day: `r medianSteps`

## What is the average daily activity pattern?

Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
library(lubridate)
library(ggplot2)

meanActivityByInterval <- activity[,.(meanSteps=mean(steps,na.rm=T)),by=.(interval)]

ggplot(meanActivityByInterval, aes(x=interval,y=meanSteps)) + 
	geom_line(stat="identity", position = 'identity')
```

On average, the maximum number of steps are made in interval: `r meanActivityByInterval[order(-meanSteps)][1]$interval`


## Imputing missing values
```{r missingvals}
missingValuesCount <- activity[is.na(steps),.N]

#strategy to imput data is to get median steps per interval (where steps != NA)
# then replace NAs by that median for given interval
medianStepsByInterval <- activity[!is.na(steps),.(steps=median(steps)),by=interval]
setkey(medianStepsByInterval,interval)
setkey(activity,interval)
activity <- medianStepsByInterval[activity]
activity[!is.na(i.steps),steps:=i.steps]
activity$i.steps <- NULL #this just removes i.steps column

stepsPerDay <- activity[,.(steps=sum(steps)),by=date]
totalStepsImput <- sum(stepsPerDay$steps)
meanStepsImput <- mean(stepsPerDay$steps)
medianStepsImput <- median(stepsPerDay$steps)
hist(stepsPerDay$steps)
```

Number of missing values: `r missingValuesCount`

Mean number of steps per day: `r meanStepsImput`

Median number of steps per day: `r medianStepsImput`

###Difference before and after imputing data

Difference between imputed mean and raw mean: `r meanStepsImput-meanSteps`

Difference between imputed median and raw median: `r medianStepsImput-medianSteps`

Difference between imputed total steps and raw total steps: `r totalStepsImput-totalSteps`

## Are there differences in activity patterns between weekdays and weekends?

Panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
library(lubridate)
library(ggplot2)
activity[,weekday:=ifelse(wday(date)<3,"Weekend","Weekday")]
#hist(log(activity[weekday %in% 3:7]$steps))

ggplot(activity[,.(meanSteps=mean(steps)),by=.(interval,weekday)], aes(x=interval,y=meanSteps)) + 
	geom_line(stat="identity", position = 'identity') + facet_grid(weekday ~ .)

```

