---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(data.table)
unzip("./activity.zip")
activity <- fread("./activity.csv")

#add datetime column combining date and interval
#activity[,datetime:=ymd_hms(paste(date,interval%/%100,interval%%100,"00",sep="-"))]
```



## What is mean total number of steps taken per day?
```{r results='hide'}
stepsPerDay <- activity[!is.na(steps)][,.(steps=sum(steps)),by=date]
totalSteps <- sum(stepsPerDay$steps)
meanSteps <- mean(stepsPerDay$steps)
medianSteps <- median(stepsPerDay$steps)
hist(stepsPerDay$steps)
```

Total number of steps: `r totalSteps`

Mean number of steps per day: `r meanSteps`

Median number of steps per day: `r medianSteps`

## What is the average daily activity pattern?

Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

To get rid of the "saw effect", intervals are converted to POSIXit dates, using today's date for the date part (as the date itself does not play the role) and intervals for times.

```{r}
library(lubridate)
library(ggplot2)

meanActivityByInterval <- activity[,.(meanSteps=mean(steps,na.rm=T)),by=.(interval)]
meanActivityByInterval[,dateinterval:=ymd_hms(paste(today(),interval%/%100,interval%%100,"00",sep="-"))]

ggplot(meanActivityByInterval, aes(x=dateinterval,y=meanSteps)) + 
	geom_line(stat="identity") 
```

On average, the maximum number of steps are made in interval: `r meanActivityByInterval[order(-meanSteps)][1]$interval`


## Imputing missing values
```{r missingvals}
missingValuesCount <- activity[is.na(steps),.N]

#strategy to imput data is to get median steps per interval (where steps != NA)
# then replace NAs by that median for given interval
medianStepsByInterval <- activity[!is.na(steps),.(steps=median(steps)),by=interval]
setkey(medianStepsByInterval,interval)
setkey(activity,interval)
activity <- medianStepsByInterval[activity]
activity[!is.na(i.steps),steps:=i.steps]
activity$i.steps <- NULL #this just removes i.steps column

stepsPerDay <- activity[,.(steps=sum(steps)),by=date]
totalStepsImput <- sum(stepsPerDay$steps)
meanStepsImput <- mean(stepsPerDay$steps)
medianStepsImput <- median(stepsPerDay$steps)
hist(stepsPerDay$steps)
```

Number of missing values: `r missingValuesCount`

Mean number of steps per day: `r meanStepsImput`

Median number of steps per day: `r medianStepsImput`

###Difference before and after imputing data

Difference between imputed mean and raw mean: `r meanStepsImput-meanSteps`

Difference between imputed median and raw median: `r medianStepsImput-medianSteps`

Difference between imputed total steps and raw total steps: `r totalStepsImput-totalSteps`

## Are there differences in activity patterns between weekdays and weekends?

Panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 
```{r}
library(lubridate)
library(ggplot2)
activity[,weekday:=ifelse(wday(date)<3,"Weekend","Weekday")]
stepsPerIntervalWeekday <- activity[,.(meanSteps=mean(steps,na.rm=T)),by=.(interval,weekday)]
stepsPerIntervalWeekday[,dateinterval:=ymd_hms(paste(today(),interval%/%100,interval%%100,"00",sep="-"))]

ggplot(stepsPerIntervalWeekday, aes(x=dateinterval,y=meanSteps)) + 
	geom_line(stat="identity", position = 'identity') + facet_grid(weekday ~ .)

```

