---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
library(data.table)
unzip("./activity.zip")
activity <- fread("./activity.csv")
```



## What is mean total number of steps taken per day?
```{r results='hide'}
stepsPerDay <- activity[!is.na(steps)][,.(steps=sum(steps)),by=date]
totalSteps <- sum(stepsPerDay$steps)
meanSteps <- mean(stepsPerDay$steps)
medianSteps <- median(stepsPerDay$steps)
```

Total number of steps: `r totalSteps`

Mean number of steps per day: `r meanSteps`

Median number of steps per day: `r medianSteps`

## What is the average daily activity pattern?



## Imputing missing values
```{r missingvals}
missingValues <- activity[is.na(steps),.N]
medianSteps <- activity[!is.na(steps),.(steps=median(steps)),by=interval]
setkey(medianSteps,interval)
setkey(activity,interval)
imput <- medianSteps[activity]
imput[!is.na(i.steps),steps:=i.steps]
imput$i.steps <- NULL

stepsPerDay <- imput[,.(steps=sum(steps)),by=date]
totalSteps <- sum(stepsPerDay$steps)
meanSteps <- mean(stepsPerDay$steps)
medianSteps <- median(stepsPerDay$steps)

hist(stepsPerDay$steps)
```

Number of missing values: `r missingValues`

Mean number of steps per day: `r meanSteps`

Median number of steps per day: `r medianSteps`

## Are there differences in activity patterns between weekdays and weekends?
